---
title: "Mapletech Data Exploration"
author: "Karen Jiang"
date: "4/6/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(readxl)
library(janitor)
library(ggthemes)
library(ggalt)

results <- read_excel(path = "data/Game-Results.xlsx") %>%
  clean_names() %>%
  mutate(
    target_price = as.integer(target_price),
    at_what_price_per_unit = as.numeric(at_what_price_per_unit),
    what_was_the_first_offer = as.numeric(what_was_the_first_offer),
    what_was_the_counteroffer = as.numeric(what_was_the_counteroffer)
  ) %>%
  mutate(did_you_reach_an_agreement = replace(did_you_reach_an_agreement,
                                              did_you_reach_an_agreement == "-", 
                                              "No"),
         role = as.factor(role),
         diff_first_counter = abs(what_was_the_first_offer - what_was_the_counteroffer),
         diff_first_final = abs(what_was_the_first_offer - at_what_price_per_unit))


zopa = c(10, 35)
```

## Strike a Deal!

Only one group in our class did not reach an agreement on a final price for this exercise. 

Group 48, are you real data? Regardless, we'll filter you out from the following analysis.
```{r, fig.width=5, fig.align='center'}
results %>% 
  count(did_you_reach_an_agreement) %>%
  ggplot(aes(x = "", y = n, fill = did_you_reach_an_agreement)) + 
  geom_bar(stat = "identity", width = 1, color="white") +
  coord_polar("y", start = 0) + 
  theme_void() +
  labs(
    fill = "Did You Reach an Agreement?"
  )
```

## BATNA and Who Goes First?

The **BATNA** is the *Best Alternative to a Negotiated Agreement*.

Mapletech's BATNA would have been no deal (no gains, no losses).
Yazawa's BATNA on the other hand would have been to buy the same part from another manufacturer at $35 / unit.


Mapletech might have had the upper hand in this negotiation, but they also more frequently made the first offer. 

```{r fig.width = 5, fig.align = "center"}
results %>% 
  count(who_made_the_first_offer) %>%
  filter(who_made_the_first_offer != "-") %>%
  ggplot(aes(x = "", y = n, fill = who_made_the_first_offer)) + 
  geom_bar(stat = "identity", width = 1, color="white") +
  coord_polar("y", start = 0) + 
  theme_void() +
  labs(
    fill = "Who made the First Offer?"
  )
```

## Zone of Possible Agreement

Mapletech's reservation point was $10.
Yazawa's reservation point was $35.

Any amount between that is the **Zone of Possible Agreement**, or **ZOPA**.


## Target Price

Target prices were generally anchored around the values provided by the case guidelines. 
However, there are a handful of folks in Mapletech who were less constrained by anchors, and chose even higher target prices. 


```{r}
results %>%
  filter(!is.na(target_price),
         group != 10) %>%
  ggplot(aes(x = target_price)) +
  annotate("rect", xmin = 10, xmax = 35, ymin = 0, ymax = Inf, fill = "gray", alpha = 0.2) +
  annotate("text", x = 22.5 , y = 0.05, label = "ZOPA") + 
  geom_vline(xintercept = zopa, linetype = "dotted") +
  geom_density(aes(fill = role), alpha = 0.5) +
  theme_classic() +
  labs(
    title = "What was the Target Price for Mapletech vs. Yazawa?",
    x = "$ / Unit",
    y = "Density", 
    fill = "Role"
  )
  
```

How does an internal, unstated target price affect the First Offer?

## First Offer

Yazawa's first offers were near $10 range (Mapletech's reservation point), right at the edge of the ZOPA, while
Mapletech's first offers were around $40, outside of the ZOPA.

The range of first offers is more extreme than the range of target prices for either group. Since Mapletech is not bounded by an upper limit, we see that they


```{r}
results %>%
  filter(!is.na(target_price),
         group != 10) %>%
  ggplot(aes(x = what_was_the_first_offer)) +
  annotate("rect", xmin = 10, xmax = 35, ymin = 0, ymax = Inf, fill = "gray", alpha = 0.2) +
  annotate("text", x = 22.5 , y = 0.065, label = "ZOPA") + 
  geom_vline(xintercept = zopa, linetype = "dotted") +
  geom_density(aes(fill = who_made_the_first_offer), alpha = 0.5) +
  theme_classic() +
  labs(
    title = "What was the First Offer made by Mapletech vs. Yazawa?",
    x = "$ / Unit",
    y = "Density", 
    fill = "Role"
  )
```



## Did making the First Offer make a difference?

The results are counter-intuitive. 

We would have expected that if Mapletech made the first offer and set an high anchor, we would expect that a peak in the final price near the higher end of the ZOPA. Instead, we see a relatively uniform distribution between %15 and $30.

When Yazawa made the first offer, there is a bimodal distribution with peaks at $20 and $30, suggesting that Yazawa making the first offer was only effective in lowering the final agreed price for a few cases.

However almost all deals were reached within the ZOPA (as should be the case).

```{r}
results %>%
  filter(did_you_reach_an_agreement == "Yes") %>%
  ggplot(aes(x = at_what_price_per_unit)) +
  annotate("rect", xmin = 10, xmax = 35, ymin = 0, ymax = Inf, fill = "gray", alpha = 0.2) +
  annotate("text", x = 22.5 , y = 0.06, label = "ZOPA") +
  geom_vline(xintercept = zopa, linetype = "dotted") +
  geom_density(alpha = 0, linetype = "dashed") + 
  geom_density(aes(fill = who_made_the_first_offer), alpha = 0.4) +
  theme_classic() +
  labs(
    title = "What was the Final Price Agreed Upon by both parties?",
    x = "Final Price per Unit",
    y = "Density",
    fill = "First Offer made by: "
  ) +
  theme(
    legend.position = "top"
  )
```

How might this result be explained by the negotiation process?

## The Negotiation

### First Offer and Counter Offer

First, let's understand what's happening with the visual: 

- **Black Dots** represent the amount of the First Offer.
- <font color = "darkorange">Orange Dots</font> represent the Counter Offer.
- The graph is split by who made the First Offer 
  - Mapletech on the Left
  - Yazawa on the Right

Generally, when the first offer was an extreme low/high ball, they were met with an aggressive counter. 
The counter offer range tends to decrease if the first offer was closer to center of the ZOPA.

Group 37 represents a case of a very reasonable First Offer that was accepted immediately. 

<br>
<br>
```{r}
results %>%
  filter(role == "Yazawa",
         !is.na(target_price)) %>%
 ggplot(aes(x = what_was_the_first_offer, 
            xend = what_was_the_counteroffer, 
            y = reorder(group, what_was_the_first_offer))
        ) + 
  annotate("rect", 
           xmin = 10, xmax = 35, ymin = -Inf, ymax = Inf, 
           fill = "gray", 
           alpha = 0.2) +
  geom_vline(xintercept = zopa, 
             linetype = "dotted") +
  facet_grid(
    cols = vars(who_made_the_first_offer) 
  ) + 
  geom_dumbbell(
    colour_xend = "darkorange",
    size_x = 2,
    size_xend = 1
  ) + 
  theme_classic() +
  labs(
    title = "How did Groups Respond to the First Offer with a Counter Offer?",
    subtitle = "based on who made the First Offer",
    x = "$ / Unit",
    y = "Group"
  ) +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    axis.text.y.left = element_text(size = 6)
  )
```


### Final Amount

Now we can add the Final amount to the graph, which is represented by a <font color = "blue">Blue X</font>.

The final offers tend to split the difference between the First Offer and Counter Offer.

```{r}
results %>%
  filter(role == "Yazawa",
         !is.na(target_price)) %>%
 ggplot(aes(x = what_was_the_first_offer, 
            xend = what_was_the_counteroffer, 
            y = reorder(group, what_was_the_first_offer))
        ) + 
  annotate("rect", 
           xmin = 10, xmax = 35, ymin = -Inf, ymax = Inf, 
           fill = "gray", 
           alpha = 0.2) +
  geom_vline(xintercept = zopa, 
             linetype = "dotted") +
  facet_grid(
    cols = vars(who_made_the_first_offer) 
  ) + 
  geom_dumbbell(
    colour_xend = "darkorange",
    size_x = 2,
    size_xend = 1
  ) + 
  geom_point(
    aes(x = at_what_price_per_unit), 
    shape = 4, 
    size = 1, 
    color = "blue"
  ) + 
  theme_classic() +
  labs(
    title = "How did Groups Respond to the First Offer with a Counter Offer?",
    subtitle = "based on who made the First Offer",
    x = "$ / Unit",
    y = "Group"
  ) +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    axis.text.y.left = element_text(size = 6)
  )
  
``` 
In certain cases, a good counter-offer can be just as effective in setting an achor as the First Offer.



```

